### 2017-11-06 acat_paper/paper.tex
Added pdf for acat from xelatex

\documentclass[a4paper]{jpconf}
\usepackage{graphicx}
\begin{document}
\title{Some Title}

\author{Egor Khairullin, Andrey Ustuzhanin}

% \address{Production Editor, \jpcs, \iopp, Dirac House, Temple Back, Bristol BS1~6BE, UK}

\ead{mikari@yandex-team.ru, anaderi@yandex-team.ru}

\begin{abstract}
The result of many machine learning algorithms are computational complex models. And further growth in the quality of the such models usually leads to a deterioration in the applying times. However, such high quality models are desirable to be used in the conditions of limited resources (memory or cpu time).
This article discusses how to trade the quality of the model for the speed of its applying a novel boosted trees algorithm called Catboost. The idea is to combine two approaches: training fewer trees and uniting trees into huge cubes. The proposed method allows for pareto-optimal reduction of the computational complexity of the decision tree model with regard to the quality of the model. In the considered example number of lookups was decreased from 5000 to only 6 (speedup factor of 1000) while AUC score of the model was reduced by less than per mil.
\end{abstract}

\section{Introduction}
Today, Machine Learning (ML) is used in various fields ranging from information search in the Internet and speech recognition to optimization of the steel composition and the search for new physics.
In almost all cases, resources (memory, CPU, network, etc.) are very limited to solve the task.

For the most ML algorithms, improving the quality of the model usually leads to higher levels of resource consumption. For example, with the use tree-based boosting algorithms, increasing the number of trees from tens to thousands usually leads to a significant increase in quality, while increasing the load on memory and CPU when applying such a model. Increased requirements on the hardware resources of more accurate models is usually solved by various technical optimizations of storage and application of the resulting model: for example, using vector calculations. Another approach is the modification of the learning process, withch makes it possible to obtain models with certain properties that make it possible to signigicantly accelerate their application without maintaining the required level of quality.

In this paper, we consider the modification of the learning process of predictive models by the algorithm CatBoost [1], which will accelerate the apply of the model while maintaining the required level of quality.

\section{CatBoost}

CatBoost is the implementation of a tree-based gradient boosting in which so-caled oblivious decision trees are used. In this algorithm the values of each feature $i$ are discretized into baskets using the boundaries $B_i$. The boundaries themselves are determined in advance using various statistics.
Source vector with real values is replaced by a vector binary values $0$ and $1$ ($false$ and $true$). The value of the $i$-th attribute $f_i$ is replaced by the binary vector $g_i$ where $g_{ij} = f_i > B_{ij}$. Finally, all $g_i$ are concatenated into one vector. Consecutively, in the resuling trees instead of comparing some real feature value with some threshold there will be just boolean check.

Number of baskets could by controlled by passing the appropriate parameter during learning procedure.

Also another important parameter should be marked. It is the speed of learning. The smaller it is, the smaller the contribution of each new tree. However, this has to be compensated by a large number of trees. The speed of learning can be selected by cross-validation process with the fixed number of trees.
Usually, more trees means higher quality.

\section{Decision Tensor}
One of the simplest ways to accelerate is the preliminary calculation. The feature vector obtained after binarization takes a limited number of values. Therefore, we can precompute the prediction for all such values.

Let $n$ be the number of features. Let us consider some model $Z$, trained by some algorithm, containing a set of trees $T$. For convenience, we assume the values of the features lie in thre interval $[0, 1]$.

Let $f_i$ be the set of basket boundaries for the $i$-th feature, which was used in $T$. Then using the planes $X_i = b_{ij}$ for all $i,j$, we can split the space $[0;1]$ int small hyperparallelepipeds. It is obvious that within each hyperparallelepiped the prediction of the model is constant.
In fact, the model $Z$ allows you to calculate the value at any point of our hypercube (or Tensor). However, if we pre-calculate the value in each piece, then we will accelerate the application of our model, reducin it only determining the desired piece. The asymptotic complexity of the application of the tree-based mode it $O(h\cdot t)$ where $h$ is the height of the trees, $t$ is the total number of trees. Asymptotic complexity of Tensor is $O(n\cdot log(b))$ where $n$ is the number of features and $b$ is the number of cuts per feature (this complexity is achieved by using binary search). It should be marked for small $b$ simple linear search could be a bit faster that binary. So, with a small number of features Tensor can be applied much faster than usual tree-based model.

However, Decision Tensor has large memory footpring. It needs $O(b^n)$ of memory and it may be too large for common computer. Therefore, the precomputed tensor can only be used for small $b$ and $n$.

In the general, the size of one tensor can be defined as follows:

$$D = (f_{1}, ..., f_{n})$$

$$f_{i} = \{b_{i1}, ..., b_{ip_{i}}\}$$

$$S(D) \propto \prod |f_{i}| = \prod p_i$$

Consider the case for small n. We can get a small $b$ by two main ways: limit the number of cuts for each feature and train model with big enough tree count or severely limit the tree count, then the ottal number of really used cuts will be exactly $h\cdot t$ without taking into account repeated use of the same cuts.

The first method was used in [2]: there was trained a CatBoost with a small number of cuts and turned into a Decision Tensor.

The consumption of memory in one hypercube can be extremely huge provded that acceptable quality is achieved, so the natural continuation is the construction of several Decision Tensors.

\section{Tensor Similarity}
Introduce the metric of the similarity of two tensors. Let's define it as the difference between the sum of their sizes and the size of their union:

$$D^1 = (f^{1}_{1}, ..., f^{1}_{n})$$

$$D^2 = (f^{2}_{1}, ..., f^{2}_{n})$$

$$D^u = D^1 + D^2$$

$$D^u = (f^{u}_{1}, ..., f^{u}_{f}), f^{u}_{i} = f^{1}_{i} \cup f^{2}_{i}$$

$$Sim(D^1, D^2) = S(D^1) + S(D^2) - S(D^1 + D^2)$$

\section{Multicriteria optimization}

Target metrics are $ Q $ (model quality), $ S $ (size), and $ t $ (time of application). However, we can assume that the application time is proportional to the number of decision tensors $ M $.
The input vector will need to describe the algorithm for obtaining trees and the algorithm for combining them into cubes. We denote it by $ v $ and by $V$ the feasible set of input vectors. Then we need to solve the following problem:

$$min(Q(V), S(V), M(V)), v \in V$$

Before we move to consider the contents of the vector $ X $, it should be noted that for practical application this problem can be solved using the constraint method. Usually in practice there are quite strict limitations on the size of the model and the available application time.

\section{Algorithm of combining a particular set of trees in a set of decision tensors}

Suppose we have $N$ trees and we want to combine them in $M$ of the decision tensors. The metric of optimality of the partition will be assumed to be occupied memory.
In general, this is not an easy task and it requires a separate study. We used the following heuristic approach:
\medskip
\renewcommand{\theenumi}{\arabic{enumi}}
\begin{enumerate}
\item Initialize $M$ tensors by placing in each one an arbitrarily chosen tree
\item Choose any of the remaining trees
\item Combine it with the most similar tensor
\item Repeat steps 2-3 until there are no unused trees left.
\end{enumerate}

\section{A complete algorithm for constructing an Decision Tensors Ensemble with the help of CatBoost}

There are two key approaches:

\medskip
\begin{itemize}
\item We fix a very rough partition with a small number of cubes and train an arbitrarily large number of trees.
\item The partitioning is done quite detailed, but at the same time we significantly reduce the number of trained trees, thereby reducing the number of actually used buckets.
\end{itemize}

The algorithm actively uses both these ideas.

\medskip
\renewcommand{\theenumi}{\arabic{enumi}}
\begin{enumerate}
\item Determine the desired number of tensors (M) based on time constraints
\item Train the CatBoost model with a limited number of buckets
\item Unite the trees from the obtained model into the first decisive tensor from the ensemble
\item Use random search to sort through the key parameters of CatBoost
\medskip
\begin{itemize}
\item Train the CatBoost model, using the previously trained model as the baseline
\item Transform into M-1 Decision Tensors
\item Remove ensembles that do not fit the required size limits.
\end{itemize}
\item We choose the ensemble with the best quality on an additional test set of data.
\end{enumerate}

This algorithm allows to find the optimal Pareto ensemble of Decision Tensors.

\section{Application of the algorithm on the dataset}

We have used the dataset and the model from [@ likhomanenko2015lhcb]. Number of features in the dataset is 10, the model was trained on 5000 trees with a number of the buckets 64. We compared our algorithm for constructing Decision Tensors Ensemble with trivial approach: training a small number of trees but with high value of learning speed.
For this we divided the data set into three parts: teaching (50%), test (25%) and validation (25%). We trained the models on the training sample, then using the test we selected the best quality models under given constraints.

We have limited the number of tensor by 6 and the total size of models by one gigabyte, believing every prediction requires 4 bytes (float). We looked through only the key parameters of CatBoost about the number of trees and the speed of learning.

The area under the ROC curve where False Positive Rate less than 0.05 was taken as a metric. For convenience, we have adjusted this metric so that its maximum value (with an ideal model) is equal to 1.

\section{Discussion}

We compared the CatBoost model with 5000 trees ("best"), the fast CatBoost models for 1-125 trees ("fast") and the Decision Tensors Ensembles with different memory limits and different tensors numbers.

Chart 1 shows the dependence of quality on the number of lookups. Chart 2 shows the dependence of quality on the total time of application at the full validation dataset. Table 3 shows properties of the models trained in the process.

TODO: добавить графики из презентации

It can be seen that at a comparable quality with "fast" models, DTE allows a 25 times reduction in the number of lookups. And in comparison with the "best" model, the number of lookups is 1000 times smaller with a small loss in quality ($ <0.006 $ with the "best" quality of 0.6).

However, the difference in working time is not so significant - 2.5 times and 50 times, respectively. This is due to the principles of the cache and the specialty of our benchmark. Since the Decision Tensor takes up a very large amount of memory and does not fit into the cache because of this, almost every lookup results in reading from the RAM. CatBoost models are small enough, because of what can be entirely placed in the cache. Since our benchmark consists only of model applications, CatBoost models are not forced out of the cache. In real applications, the CatBoost model will most likely not own the cache alone and will be even slower because of this. And Decision Tensor Ensemble almost do not depend on the cache size.

\section{Conclusion}

Decision Tensor Ensbles allow you to simplify the application of complex Boosting Decision Trees algorithms (such as CatBoost) in real-time systems without the ability to use batching application.
Considered algorithm for building ensembles allows you to exchange consumed memory for the speed of application with a small loss in quality.
In addition, the use of the Ensemble can significantly improve the quality in comparison with the use of a single Decision Tensor.

\section*{References}
\numrefs{99}
\item Gulin A, Kuralenok I, Pavlov D 2011 {\it Winning The Transfer Learning Track of Yahoo!'s Learning To Rank Challenge with YetiRank.} (Yahoo! Learning to Rank Challenge) p 63--76
\item Likhomanenko T, Ilten P, Khairullin E, Rogozhnikov A, Ustyuzhanin A, Williams M 2015 {\it LHCb Topological Trigger Reoptimization} (Journal of Physics: Conference Series)
\endnumrefs

\end{document}

### 2017-11-09 acat_paper/paper.tex
first section

@@
 \begin{document}
-\title{Some Title}
+\title{Speeding up prediction performance of the BDT-based models}
 
-\author{Egor Khairullin, Andrey Ustuzhanin}
-
+\author{
+ E. Khairullin$^{1, 3}$
+ and
+ A. Ustyuzhanin$^{1, 2, 3}$
+}
 % \address{Production Editor, \jpcs, \iopp, Dirac House, Temple Back, Bristol BS1~6BE, UK}
+\address{$^{1}$ Moscow Institute of Physics and Technology}
+\address{$^{2}$ National Research University Higher School of Economics, 20 Myasnitskaya st., Moscow 101000, Russia}
+\address{$^{3}$ Yandex School of Data Analysis, 11/2, Timura Frunze st., Moscow 119021, Russia}
 
@@
 \begin{abstract}
-The result of many machine learning algorithms are computational complex models.
-And further growth in the quality of the such models usually leads to a deterioration in the applying times.
-However, such high quality models are desirable to be used in the conditions of limited resources (memory or cpu time).
-This article discusses how to trade the quality of the model for the speed of its applying a novel boosted trees algorithm called Catboost.
-The idea is to combine two approaches: training fewer trees and uniting trees into huge cubes.
+The outcome of a machine learning algorithm is a prediction model.
+It is quite usual such models are computationaly expensive and growth of the quality of such models leads to the deterioration in the inference speed.
+However it is not always tradeoff between quality vs speed.
+In this paper we show it is possible to speed up model by using additional memory without loosing prediction quality significantly for a novel boosted trees algorithm called Catboost.
+The idea is to combine two approaches: training fewer trees and merging trees into kind of hash maps (DecisionTensors).
 The proposed method allows for pareto-optimal reduction of the computational complexity of the decision tree model with regard to the quality of the model.
@@
 \section{Introduction}
-Today, Machine Learning (ML) is used in various fields ranging from information search in the Internet and speech recognition to optimization of the steel composition and the search for new physics.
-In almost all cases, resources (memory, CPU, network, etc.) are very limited to solve the task.
+Today, Machine Learning (ML) approach is used in various fields ranging from the information retrieval in the Internet and speech recognition to optimization of the steel composition and the search for the New Physics.
+The resources (memory, CPU, network, etc.) are very limited to solve the task in production settings.
 
-For the most ML algorithms, improving the quality of the model usually leads to higher levels of resource consumption.
-For example, with the use tree-based boosting algorithms, increasing the number of trees from tens to thousands usually leads to a significant increase in quality, while increasing the load on memory and CPU when applying such a model.
-Increased requirements on the hardware resources of more accurate models is usually solved by various technical optimizations of storage and application of the resulting model: for example, using vector calculations.
-Another approach is the modification of the learning process, withch makes it possible to obtain models with certain properties that make it possible to signigicantly accelerate their application without maintaining the required level of quality.
+For the most of the ML algorithms, improving the quality of the model usually leads to a higher levels of resource consumption.
+For example, with the use tree-based boosting algorithms, the increase of the number of trees from tens to thousands usually leads to the significant increase in the quality and at the same time the model size and inference time also amplify.
+The higher demands for the hardware resources for the more accurate models are usually solved by various technical optimizations model storage mechanism.
+For example vectoriation of the tree weights storage can be used.
+Another approach is the modification of the learning process that makes it possible to obtain models with certain properties that make it possible to signigicantly accelerate their application without maintaining the required level of quality.
 
-In this paper, we consider the modification of the learning process of predictive models by the algorithm CatBoost [1], which will accelerate the apply of the model while maintaining the required level of quality.
+In this paper, we consider the modification of the learning process of predictive models by the boost decision tree-based algorithm CatBoost [1].
+Such modification will accelerate the prediction time of the model while maintaining the required level of the quality.
 

### 2017-11-09 acat_paper/paper.tex
the rest

@@
 For example vectoriation of the tree weights storage can be used.
-Another approach is the modification of the learning process that makes it possible to obtain models with certain properties that make it possible to signigicantly accelerate their application without maintaining the required level of quality.
+Another approach is the modification of the training process that makes it possible to obtain models with certain properties that make it possible to signigicantly accelerate their application without maintaining the required level of quality.
 
-In this paper, we consider the modification of the learning process of predictive models by the boost decision tree-based algorithm CatBoost [1].
+In this paper, we consider the modification of the training process of a predictive models by decision tree-based algorithm CatBoost [1].
 Such modification will accelerate the prediction time of the model while maintaining the required level of the quality.
@@
 CatBoost is the implementation of a tree-based gradient boosting in which so-caled oblivious decision trees are used.
-In this algorithm the values of each feature $i$ are discretized into baskets using the boundaries $B_i$.
-The boundaries themselves are determined in advance using various statistics.
-Source vector with real values is replaced by a vector binary values $0$ and $1$ ($false$ and $true$).
-The value of the $i$-th attribute $f_i$ is replaced by the binary vector $g_i$ where $g_{ij} = f_i > B_{ij}$.
+The values of each feature $i$ are discretized into baskets using the boundaries $B_i$.
+The boundaries themselves are determined in advance using various heuristic.
+Original feature vector with real values is replaced by a vector binary values $0$ and $1$ ($false$ and $true$).
+The value of the $i$-th attribute $f_i$ is replaced by a binary vector $g_i$ where $g_{ij} = f_i > B_{ij}$.
 Finally, all $g_i$ are concatenated into one vector.
-Consecutively, in the resuling trees instead of comparing some real feature value with some threshold there will be just boolean check.
+Consecutively, the resuling decision will make bunch of boolean checks instead of comparing some real feature value with the thresholds.
 
-Number of baskets could by controlled by passing the appropriate parameter during learning procedure.
+Number of baskets used for the discretization could be controlled by specifying the appropriate parameter during the training procedure.
 
-Also another important parameter should be marked.
-It is the speed of learning.
-The smaller it is, the smaller the contribution of each new tree.
+Let's consider another important training characteristic, namely the learning rate.
+The smaller it is, the smaller the contribution of each next tree in the ensemble.
 However, this has to be compensated by a large number of trees.
-The speed of learning can be selected by cross-validation process with the fixed number of trees.
-Usually, more trees means higher quality.
+The learning rate can be chosen by the cross-validation process given the fixed number of trees.
+Usually, the more trees the higher the model quality.
 
 \section{Decision Tensor}
-One of the simplest ways to accelerate is the preliminary calculation.
-The feature vector obtained after binarization takes a limited number of values.
-Therefore, we can precompute the prediction for all such values.
+One of the simplest ways to accelerate computational performance is the preliminary calculation.
+The feature vector obtained after the discretization takes a limited number of values.
+Therefore, we can precompute the predictions for all possible values combination.
 
@@
 Let us consider some model $Z$, trained by some algorithm, containing a set of trees $T$.
-For convenience, we assume the values of the features lie in thre interval $[0, 1]$.
+For the convenience, we assume the values of every feature belongs to the interval $[0, 1]$.
 
 Let $f_i$ be the set of basket boundaries for the $i$-th feature, which was used in $T$.
-Then using the planes $X_i = b_{ij}$ for all $i,j$, we can split the space $[0;1]$ int small hyperparallelepipeds.
-It is obvious that within each hyperparallelepiped the prediction of the model is constant.
-In fact, the model $Z$ allows you to calculate the value at any point of our hypercube (or Tensor).
-However, if we pre-calculate the value in each piece, then we will accelerate the application of our model, reducin it only determining the desired piece.
-The asymptotic complexity of the application of the tree-based mode it $O(h\cdot t)$ where $h$ is the height of the trees, $t$ is the total number of trees.
-Asymptotic complexity of Tensor is $O(n\cdot log(b))$ where $n$ is the number of features and $b$ is the number of cuts per feature (this complexity is achieved by using binary search).
-It should be marked for small $b$ simple linear search could be a bit faster that binary.
-So, with a small number of features Tensor can be applied much faster than usual tree-based model.
+Then we can split the interval $[0;1]$ into small hyper-parallelepipeds.
+The prediction of the model is constant within each hyperparallelepiped.
+In fact, the model $Z$ allows you to calculate the value at any point of our hypercube that we call Tensor.
+However, if we pre-calculate the value for each piece, then we will accelerate the application of our model, effectively reducing it to the time requited for the determination of the desired piece.
+The asymptotic complexity of the application of the tree-based mode it $O(h\cdot t)$ where $h$ is the height of the trees, $t$ is the total number of the trees.
+Asymptotic complexity of Tensor is $O(n\cdot log(b))$ where $n$ is the number of the features and $b$ is the number of cuts per feature (this complexity is achieved by using binary search).
+(It should be noted for small $b$ simple linear search could be a bit faster that binary).
+So, with a small number of features the Tensor-based prediction can be computed much faster than usual tree-based model.
 
-However, Decision Tensor has large memory footpring. It needs $O(b^n)$ of memory and it may be too large for common computer. Therefore, the precomputed tensor can only be used for small $b$ and $n$.
+However, Decision Tensor has larger memory footprint.
+It needs $O(b^n)$ bytes of memory and it may be too large for common computer.
+Therefore, the precomputed tensor can only be used for small $b$ and $n$.
 
-In the general, the size of one tensor can be defined as follows:
+In the general, the size of one tensor can be estimated as follows:
 
@@
 
-Consider the case for small n.
-We can get a small $b$ by two main ways: limit the number of cuts for each feature and train model with big enough tree count or severely limit the tree count, then the ottal number of really used cuts will be exactly $h\cdot t$ without taking into account repeated use of the same cuts.
+Consider the case for small $n$.
+We can get a small $b$ in two main ways: limit the number of cuts for each feature and train the model with big enough tree count or severely limit the tree count, then the total number of effective cuts will be exactly $h\cdot t$ without taking into account repeated use of the same cuts.
 
-The first method was used in [2]: there was trained a CatBoost with a small number of cuts and turned into a Decision Tensor.
+The first method was used in [2]: there was a trained CatBoost model with a small number of cuts that has been precomputed into a Decision Tensor.
 
-The consumption of memory in one hypercube can be extremely huge provded that acceptable quality is achieved, so the natural continuation is the construction of several Decision Tensors.
+The consumption of memory for a single hypercube can be extremely huge provded that acceptable quality is achieved, so the natural extension of the idea is the construction of Decision Tensors ensemble.
 
 \section{Tensor Similarity}
-Introduce the metric of the similarity of two tensors. Let's define it as the difference between the sum of their sizes and the size of their union:
+Let's introduce the metric for the similarity of two tensors. Let's define it as the difference between the sum of their sizes and the size of their union:
 
@@
 
-Target metrics are $ Q $ (model quality), $ S $ (size), and $ t $ (time of application). However, we can assume that the application time is proportional to the number of decision tensors $ M $.
-The input vector will need to describe the algorithm for obtaining trees and the algorithm for combining them into cubes.
-We denote it by $ v $ and by $V$ the feasible set of input vectors.
+The target metrics are $ Q $ (model quality), $ S $ (size), and $ t $ (time of inference). However, we can assume that the inference time is proportional to the number of decision tensors $ M $.
+The input feature vectors will define the algorithm for obtaining the trees and the algorithm for combining them into the Tensors.
+We denote single vector by $ v $ and let $V$ be the set of all input vectors.
 Then we need to solve the following problem:
@@
 
-Before we move to consider the contents of the vector $ X $, it should be noted that for practical application this problem can be solved using the constraint method.
-Usually in practice there are quite strict limitations on the size of the model and the available application time.
+It should be noted that for practical application this problem can be solved using the constrained optimization method.
+But in practice there are quite strict limitations on the size of the model and the available application time (TODO - what did you mean?).
 
-\section{Algorithm of combining a particular set of trees in a set of decision tensors}
+\section{Algorithm for combining set of Decision Trees to a set of Decision Tensors}
 
-Suppose we have $N$ trees and we want to combine them in $M$ of the decision tensors. The metric of optimality of the partition will be assumed to be occupied memory.
+Suppose we have $N$ trees and we want to combine them in $M$ of the Decision Tensors. The efficiency metric of the partition is the result Tensor set memory footprint.
 In general, this is not an easy task and it requires a separate study. We used the following heuristic approach:
@@
 \begin{enumerate}
-\item Initialize $M$ tensors by placing in each one an arbitrarily chosen tree
+\item Initialize $M$ Tensors by placing in each one an arbitrarily chosen tree
 \item Choose any of the remaining trees
@@
 \end{enumerate}
+Such reduction of the set of trees to a set of Tensors we call partitioning. 
+(TODO: what is rough and what is detailed partition?)
 
@@
 \begin{itemize}
-\item We fix a very rough partition with a small number of cubes and train an arbitrarily large number of trees.
+\item We fix a very rough partition by a small number of Tensors and train an arbitrarily large number of trees.
 \item The partitioning is done quite detailed, but at the same time we significantly reduce the number of trained trees, thereby reducing the number of actually used buckets.
@@
 
-The algorithm actively uses both these ideas.
+The algorithm actively uses both of these ideas.
 
@@
 \begin{enumerate}
-\item Determine the desired number of tensors (M) based on time constraints
+\item Determine the desired number of Tensors (M) based on time constraints
 \item Train the CatBoost model with a limited number of buckets
-\item Unite the trees from the obtained model into the first decisive tensor from the ensemble
-\item Use random search to sort through the key parameters of CatBoost
+\item Merge the trees from the obtained model into the first decisive tensor from the ensemble
+\item Chose another set of CatBoost hyperparameters using Random Search
 \medskip
@@
 \item Train the CatBoost model, using the previously trained model as the baseline
-\item Transform into M-1 Decision Tensors
-\item Remove ensembles that do not fit the required size limits.
+\item Transform it into M-1 Decision Tensors
+\item Remove ensembles that do not fit the required size limits (TODO: specify limits at the beginning?).
 \end{itemize}
-\item We choose the ensemble with the best quality on an additional test set of data.
+\item We choose the ensemble with the best quality on an additional data sample.
 \end{enumerate}
 
-This algorithm allows to find the optimal Pareto ensemble of Decision Tensors.
+This algorithm allows to find the Pareto-optimal ensemble of Decision Tensors.
 
-\section{Application of the algorithm on the dataset}
+\section{Real problem illustration}
 
@@
 Number of features in the dataset is 10, the model was trained on 5000 trees with a number of the buckets 64.
-We compared our algorithm for constructing Decision Tensors Ensemble with trivial approach: training a small number of trees but with high value of learning speed.
-For this we divided the data set into three parts: teaching (50%), test (25%) and validation (25%).
-We trained the models on the training sample, then using the test we selected the best quality models under given constraints.
+We compared our algorithm for constructing Decision Tensors Ensemble with trivial approach: training a small number of trees but with high value of the training speed.
+For this we divided the data set into three parts: training (50\%), test (25\%) and validation (25\%).
+We trained the models on the training sample, then used the test for the selection of the best models under given constraints.
 
-We have limited the number of tensor by 6 and the total size of models by one gigabyte, believing every prediction requires 4 bytes (float).
-We looked through only the key parameters of CatBoost about the number of trees and the speed of learning.
+We have limited the number of tensor by 6 and the total size of models by one gigabyte, so every prediction requires 4 bytes (float).
+We have scanned only through the key CatBoost hyperparameters: number of trees and the learing rate.
 
-The area under the ROC curve where False Positive Rate less than 0.05 was taken as a metric.
-For convenience, we have adjusted this metric so that its maximum value (with an ideal model) is equal to 1.
+The figure of merit we've taken also from [@_likhomanenko2015lhcb], which is the area under the ROC curve where False Positive Rate less than 0.05.
+For the convenience, we have normalized this metric so that its maximum value (with an ideal model) is equal to 1.
 
@@
 
-We compared the CatBoost model with 5000 trees ("best"), the fast CatBoost models for 1-125 trees ("fast") and the Decision Tensors Ensembles with different memory limits and different tensors numbers.
+We compared the CatBoost model with 5000 trees ("best"), the fast CatBoost models for 1-125 trees ("fast") and the Decision Tensors Ensembles (DTE) with different memory limits and different tensors numbers.
 
-Chart 1 shows the dependence of quality on the number of lookups.
-Chart 2 shows the dependence of quality on the total time of application at the full validation dataset.
+Figure 1 shows the dependence of the quality on the number of lookups.
+Chart 2 shows the dependence of the quality vs the total inference time at the full validation dataset.
 Table 3 shows properties of the models trained in the process.
@@
 However, the difference in working time is not so significant - 2.5 times and 50 times, respectively.
-This is due to the principles of the cache and the specialty of our benchmark.
-Since the Decision Tensor takes up a very large amount of memory and does not fit into the cache because of this, almost every lookup results in reading from the RAM.
-CatBoost models are small enough, because of what can be entirely placed in the cache.
-Since our benchmark consists only of model applications, CatBoost models are not forced out of the cache.
-In real applications, the CatBoost model will most likely not own the cache alone and will be even slower because of this.
-And Decision Tensor Ensemble almost do not depend on the cache size.
+This is due to the principles of the CPU caching and the size of our dataset.
+Since the Decision Tensor takes up a very large amount of memory and does not fit into the CPU cache because, almost every lookup results in reading from the RAM.
+CatBoost models are small enough to fit into the cache.
+Since our benchmark consists only of model applications, CatBoost models are not moved out of the cache.
+In real applications, the CatBoost model will most likely not sit in the cache alone and will be even slower because of this.
+On the contrary the performance of Decision Tensor Ensemble will not degrade due to that fact.
 
@@
 
-Decision Tensor Ensbles allow you to simplify the application of complex Boosting Decision Trees algorithms (such as CatBoost) in real-time systems without the ability to use batching application.
-Considered algorithm for building ensembles allows you to exchange consumed memory for the speed of application with a small loss in quality.
-In addition, the use of the Ensemble can significantly improve the quality in comparison with the use of a single Decision Tensor.
+Decision Tensor Ensbles allow you to speed up the inference by complex Boosting Decision Trees algorithms (such as CatBoost) in real-time systems.
+Considered algorithm for building Tensor ensembles allows for additional tradeoff between memory and the inference speed without significant loss of the model quality.
+In addition, the use of the Ensemble can significantly improve the model quality in comparison with a single Decision Tensor.
 

### 2017-11-09 acat_paper/paper.tex
Added graphics

@@
 
-We have used the dataset and the model from [@ likhomanenko2015lhcb].
+We have used the dataset and the model from [2].
 Number of features in the dataset is 10, the model was trained on 5000 trees with a number of the buckets 64.
@@
 
-The figure of merit we've taken also from [@_likhomanenko2015lhcb], which is the area under the ROC curve where False Positive Rate less than 0.05.
+The figure of merit we've taken also from [2], which is the area under the ROC curve where False Positive Rate less than 0.05.
 For the convenience, we have normalized this metric so that its maximum value (with an ideal model) is equal to 1.
@@
 Figure 1 shows the dependence of the quality on the number of lookups.
-Chart 2 shows the dependence of the quality vs the total inference time at the full validation dataset.
-Table 3 shows properties of the models trained in the process.
+Figure 2 shows the dependence of the quality vs the total inference time at the full validation dataset.
+Table 1 shows properties of the models trained in the process.
 
-TODO: добавить графики из презентации
+\begin{figure}
+\begin{center}
+\includegraphics[width=0.8\textwidth]{"lookups.png"}
+\caption{Lookups count versus FOM for different models sizes}
+\end{center}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=0.8\textwidth]{"times.png"}
+\caption{Application time versus FOM for different models sizes}
+\end{center}
+\end{figure}
+
+\begin{table}[h]
+\caption{\label{tabone}Properties of the some models trained in the process}
+
+\begin{center}
+\lineup
+\begin{tabular}{*{4}{l}}
+\br
+Model & Roc & Time, ms & Size, Mb \cr
+\mr
+catboost-5000 & $0.6004$ & $10000$ & $<5$ \cr
+catboost-100 & $0.589$ & $160$ & $<5$ \cr
+catboost-75 & $0.5808$ & $125$ & $<5$ \cr
+catboost-10 & $0.5336$ & $30$ & $<5$ \cr
+big dte-6 & $0.5952$ & $180$ & $1000$ \cr
+big dte-3 & $0.5938$ & $145$ & $1000$ \cr
+big dte-2 & $0.5844$ & $70$ & $1000$ \cr
+small dte-2 & $0.5794$ & $55$ & $150$ \cr
+dte-1 & $0.5618$ & $30$ & $150$ \cr
+\br
+\end{tabular}
+\end{center}
+\end{table}
 

### 2017-11-10 acat_paper/paper.tex
polish1

@@
 This is due to the principles of the CPU caching and the size of our dataset.
-Since the Decision Tensor takes up a very large amount of memory and does not fit into the CPU cache because, almost every lookup results in reading from the RAM.
+Since the Decision Tensor occupies significant amount of memory and does not fit into the CPU cache almost every lookup results in (slow) reading from the RAM.
 CatBoost models are small enough to fit into the cache.
-Since our benchmark consists only of model applications, CatBoost models are not moved out of the cache.
-In real applications, the CatBoost model will most likely not sit in the cache alone and will be even slower because of this.
-On the contrary the performance of Decision Tensor Ensemble will not degrade due to that fact.
+There was a single operation type of model application in our tests, so the CatBoost models stayed persistently in the cache.
+In real applications, CatBoost model will most likely not stay in the cache alone and will be even slower because of this.
+On the contrary the performance of the Decision Tensor Ensemble will not degrade due to that fact.
 

### 2017-11-10 acat_paper/paper.tex
Made some fixes

--- 2017-11-09 acat_paper/paper.tex
@@
 Today, Machine Learning (ML) approach is used in various fields ranging from the information retrieval in the Internet and speech recognition to optimization of the steel composition and the search for the New Physics.
-The resources (memory, CPU, network, etc.) are very limited to solve the task in production settings.
+The resources (memory, CPU, network etc.) are very limited to solve the task in production settings.
 
 For the most of the ML algorithms, improving the quality of the model usually leads to a higher levels of resource consumption.
-For example, with the use tree-based boosting algorithms, the increase of the number of trees from tens to thousands usually leads to the significant increase in the quality and at the same time the model size and inference time also amplify.
-The higher demands for the hardware resources for the more accurate models are usually solved by various technical optimizations model storage mechanism.
-For example vectoriation of the tree weights storage can be used.
-Another approach is the modification of the training process that makes it possible to obtain models with certain properties that make it possible to signigicantly accelerate their application without maintaining the required level of quality.
+For example, when using tree-based boosting algorithms, the increase of the number of trees from tens to thousands usually leads to significant increase in the quality and at the same time the model size and inference time also grow.
+Higher demands for the computational resources of the more accurate models are usually solved by various technical optimizations model storage mechanism.
+For example vectorisation of the tree weights storage can be used.
+Another approach is the modification of the training process that makes it possible to obtain models with certain properties that will provide acceleration to their application without maintaining the required level of quality.
 
-In this paper, we consider the modification of the training process of a predictive models by decision tree-based algorithm CatBoost [1].
-Such modification will accelerate the prediction time of the model while maintaining the required level of the quality.
-
+In this paper, we consider a modification of the training process of a predictive models by decision tree-based algorithm CatBoost [1] (TODO: 13).
 \section{CatBoost}
 
-CatBoost is the implementation of a tree-based gradient boosting in which so-caled oblivious decision trees are used.
-The values of each feature $i$ are discretized into baskets using the boundaries $B_i$.
-The boundaries themselves are determined in advance using various heuristic.
-Original feature vector with real values is replaced by a vector binary values $0$ and $1$ ($false$ and $true$).
-The value of the $i$-th attribute $f_i$ is replaced by a binary vector $g_i$ where $g_{ij} = f_i > B_{ij}$.
-Finally, all $g_i$ are concatenated into one vector.
-Consecutively, the resuling decision will make bunch of boolean checks instead of comparing some real feature value with the thresholds.
+CatBoost is the implementation of a tree-based gradient boosting in which oblivious decision trees are used.
+The values of each feature $i$ are discretized into baskets using the boundaries $B_i={b_{i1},...,b_{ip_i}}$, where $p_i$ is the number of boundaries.
+The boundaries themselves are determined in advance using various approaches.
+Original feature vector with real values is replaced by a vector of binary values $0$ and $1$ ($false$ and $true$).
+The value of the $i$-th attribute $f_i$ is replaced by a binary vector $G_i={g_{i1},...,g_{ip_i}}$ where $g_{ij} = f_i > b_{ij}$.
+Finally, all $G_i$ are concatenated into one vector.
+Consecutively, the resuling decision will a make bunch of boolean checks instead of comparing some real feature value with the thresholds.
 
@@
 Usually, the more trees the higher the model quality.
+(TODO: add conclusion)
 
@@
 Let us consider some model $Z$, trained by some algorithm, containing a set of trees $T$.
-For the convenience, we assume the values of every feature belongs to the interval $[0, 1]$.
+For convenience, we assume that the values of every feature belongs to the interval $[0;1]$.
+In other words, every feature vectors lies in n-dimensional cube $[0;1]$.
 
-Let $f_i$ be the set of basket boundaries for the $i$-th feature, which was used in $T$.
-Then we can split the interval $[0;1]$ into small hyper-parallelepipeds.
+Let $B_i$ be the a of basket boundaries for the $i$-th feature, which was used in $T$.
+Then we can split the hyper-cube $[0;1]$ into small hyper-parallelepipeds.
 The prediction of the model is constant within each hyperparallelepiped.
-In fact, the model $Z$ allows you to calculate the value at any point of our hypercube that we call Tensor.
-However, if we pre-calculate the value for each piece, then we will accelerate the application of our model, effectively reducing it to the time requited for the determination of the desired piece.
-The asymptotic complexity of the application of the tree-based mode it $O(h\cdot t)$ where $h$ is the height of the trees, $t$ is the total number of the trees.
-Asymptotic complexity of Tensor is $O(n\cdot log(b))$ where $n$ is the number of the features and $b$ is the number of cuts per feature (this complexity is achieved by using binary search).
-(It should be noted for small $b$ simple linear search could be a bit faster that binary).
-So, with a small number of features the Tensor-based prediction can be computed much faster than usual tree-based model.
+In fact, the model $Z$ allows you to calculate the value at any point of our hypercube.
+if we pre-calculate the value for each piece, then we will accelerate the application of our model, effectively reducing it to the time required to search a piece that contains the point.
+We called such approach a Decision Tensor.
+The asymptotic complexity of the application of the tree-based mode it $\mathcal(h\cdot t)$ where $h$ is the height of the trees, $t$ is the total number of the trees.
+Asymptotic complexity of Tensor is $\mathcal(n\cdot \log(b))$ where $n$ is the number of the features and $b$ is the number of cuts per feature (this complexity is estimated by using binary search).
+It should be noted for small $b$ simple linear search could be a bit faster that binary.
+So, with a small number of features the Tensor-based prediction can be computed much faster than usual tree-based model (TODO: 37)
 
@@
 
-In the general, the size of one tensor can be estimated as follows:
+In the general, the size of one tensor $D$ can be estimated as follows:
 
-$$D = (f_{1}, ..., f_{n})$$
+$$D = (B_{1}, ..., B_{n}),$$
 
-$$f_{i} = \{b_{i1}, ..., b_{ip_{i}}\}$$
+$$B_{i} = \{b_{i1}, ..., b_{ip_{i}}\},$$
 
-$$S(D) \propto \prod |f_{i}| = \prod p_i$$
+$$S(D) \propto \prod |B_{i}| = \prod p_i$$
 
@@
 
-The first method was used in [2]: there was a trained CatBoost model with a small number of cuts that has been precomputed into a Decision Tensor.
+The first method was used in [2]: there a trained MatrixNet (ancestor of CatBoost) model with a small number of cuts that has been precomputed into a Decision Tensor.
 
-The consumption of memory for a single hypercube can be extremely huge provded that acceptable quality is achieved, so the natural extension of the idea is the construction of Decision Tensors ensemble.
+The consumption of memory for a single hypercube can be extremely huge provided that acceptable quality is achieved, so the natural extension of the idea is the construction of Decision Tensors ensemble.
+(TODO: 45)
 
 \section{Tensor Similarity}
-Let's introduce the metric for the similarity of two tensors. Let's define it as the difference between the sum of their sizes and the size of their union:
-
-$$D^1 = (f^{1}_{1}, ..., f^{1}_{n})$$
-
-$$D^2 = (f^{2}_{1}, ..., f^{2}_{n})$$
-
-$$D^u = D^1 + D^2$$
-
-$$D^u = (f^{u}_{1}, ..., f^{u}_{f}), f^{u}_{i} = f^{1}_{i} \cup f^{2}_{i}$$
+Let's introduce the metric for the similarity of two tensors.
+Suppose, we have two tensors: $D^1 = (B^{1}_{1}, ..., B^{1}_{n})$ and $D^2 = (B^{2}_{1}, ..., B^{2}_{n})$.
+Their sum $D^u = D^1 + D^2$ could be defined as $D^u = (f^{u}_{1}, ..., f^{u}_{f})$, where $f^{u}_{i} = f^{1}_{i} \cup f^{2}_{i}$.
+Now, let's define similarity as the difference between the sum of their sizes and the size of their union:
 
@@
 \section{Multicriteria optimization}
-
+(TODO - probably should be removed)
 The target metrics are $ Q $ (model quality), $ S $ (size), and $ t $ (time of inference). However, we can assume that the inference time is proportional to the number of decision tensors $ M $.
@@
 
-Suppose we have $N$ trees and we want to combine them in $M$ of the Decision Tensors. The efficiency metric of the partition is the result Tensor set memory footprint.
+Suppose we have $T$ trees and we want to combine them in $M$ Decision Tensors. The efficiency metric of the partition is the result Tensor set memory footprint.
 In general, this is not an easy task and it requires a separate study. We used the following heuristic approach:
@@
 Such reduction of the set of trees to a set of Tensors we call partitioning.
-(TODO: what is rough and what is detailed partition?)
 
-\section{A complete algorithm for constructing an Decision Tensors Ensemble with the help of CatBoost}
+\section{A complete algorithm for constructing an Decision Tensors Ensemble}
 
@@
 \begin{itemize}
-\item We fix a very rough partition by a small number of Tensors and train an arbitrarily large number of trees.
+\item We fix a very rough (with very few ) partition by a small number of Tensors and train an arbitrarily large number of trees.
 \item The partitioning is done quite detailed, but at the same time we significantly reduce the number of trained trees, thereby reducing the number of actually used buckets.
@@
 \begin{enumerate}
+\item Determine time and size constraints
 \item Determine the desired number of Tensors (M) based on time constraints
-\item Train the CatBoost model with a limited number of buckets
-\item Merge the trees from the obtained model into the first decisive tensor from the ensemble
-\item Chose another set of CatBoost hyperparameters using Random Search
+\item Train the model with a limited number of buckets
+\item Merge the trees from the obtained model into the first Decision Tensor from the ensemble
+\item Choose another set of hyperparameters using random search
 \medskip
 \begin{itemize}
-\item Train the CatBoost model, using the previously trained model as the baseline
+\item Train the model, using the previously trained model as the baseline
 \item Transform it into M-1 Decision Tensors
-\item Remove ensembles that do not fit the required size limits (TODO: specify limits at the beginning?).
 \end{itemize}
-\item We choose the ensemble with the best quality on an additional data sample.
+\item Remove ensembles that do not fit the required size limits
+\item Choose the ensemble with the best quality on an additional data sample.
 \end{enumerate}
@@
 
-We have used the dataset and the model from [2].
-Number of features in the dataset is 10, the model was trained on 5000 trees with a number of the buckets 64.
+We have used CatBoost as a main algorithm and also took the dataset and the model from [2].
+Number of features in the dataset is 10, the model was trained with 5000 trees and 64 buckets.
 We compared our algorithm for constructing Decision Tensors Ensemble with trivial approach: training a small number of trees but with high value of the training speed.
@@
 
-We compared the CatBoost model with 5000 trees ("best"), the fast CatBoost models for 1-125 trees ("fast") and the Decision Tensors Ensembles (DTE) with different memory limits and different tensors numbers.
+We compared the CatBoost model with 5000 trees, the fast CatBoost models for 1-125 trees and the Decision Tensors Ensembles (DTE) with different memory limits and different tensors numbers.
 
@@
 
-\begin{figure}
+\begin{figure}[h]
 \begin{center}
@@
 
-It can be seen that at a comparable quality with "fast" models, DTE allows a 25 times reduction in the number of lookups.
-And in comparison with the "best" model, the number of lookups is 1000 times smaller with a small loss in quality ($ <0.006 $ with the "best" quality of 0.6).
+It can be seen that at a comparable quality with fast CatBoost models (1-125 trees), DTE allows a 25 times reduction in the number of lookups.
+And in comparison with the best model (5000 trees), the number of lookups is 1000 times smaller with a small loss in quality ($ <0.006 $ with the best quality of 0.6).
 
@@
 Considered algorithm for building Tensor ensembles allows for additional tradeoff between memory and the inference speed without significant loss of the model quality.
-In addition, the use of the Ensemble can significantly improve the model quality in comparison with a single Decision Tensor.
+In addition, the use of the Ensemble can significantly improve the model quality in comparison with a single Decision Tensor and can be used in the triggers.
 

### 2017-11-11 acat_paper/paper.tex
Fixed TODO, removed paragraph about multicriteria optimisation

@@
 
-In this paper, we consider a modification of the training process of a predictive models by decision tree-based algorithm CatBoost [1] (TODO: 13).
+In this paper, we consider a modification of the training process of a predictive models by decision tree-based algorithm CatBoost [1].
+But our approach is applicable to other boosting decision tree-based algorithms.
+
 \section{CatBoost}
@@
 The smaller it is, the smaller the contribution of each next tree in the ensemble.
-However, this has to be compensated by a large number of trees.
+However, this has to be compensated by a large number of trees and vice versa.
 The learning rate can be chosen by the cross-validation process given the fixed number of trees.
 Usually, the more trees the higher the model quality.
-(TODO: add conclusion)
+
+With default parameters CatBoost shows already good enough quality, so we variate only number of trees (and learning rate to compensate) because number of trees affects time.
+For other parameters was choosen default values.
 
@@
 It should be noted for small $b$ simple linear search could be a bit faster that binary.
-So, with a small number of features the Tensor-based prediction can be computed much faster than usual tree-based model (TODO: 37)
+So, with a small number of features the Tensor-based prediction may be computed much faster than tree-based model.
 
@@
 
-The consumption of memory for a single hypercube can be extremely huge provided that acceptable quality is achieved, so the natural extension of the idea is the construction of Decision Tensors ensemble.
-(TODO: 45)
+The consumption of memory for a single tensor can be extremely huge provided that acceptable quality is achieved, so the natural extension of the idea is the construction of Decision Tensors ensemble but with some deterioration of applying time.
 
@@
 
-\section{Multicriteria optimization}
-(TODO - probably should be removed)
-The target metrics are $ Q $ (model quality), $ S $ (size), and $ t $ (time of inference). However, we can assume that the inference time is proportional to the number of decision tensors $ M $.
-The input feature vectors will define the algorithm for obtaining the trees and the algorithm for combining them into the Tensors.
+% \section{Multicriteria optimization}
+% (TODO - probably should be removed)
+% The target metrics are $ Q $ (model quality), $ S $ (size), and $ t $ (time of inference). However, we can assume that the inference time is proportional to the number of decision tensors $ M $.
+% The input feature vectors will define the algorithm for obtaining the trees and the algorithm for combining them into the Tensors.
 We denote single vector by $ v $ and let $V$ be the set of all input vectors.
@@
 
-$$min(Q(V), S(V), M(V)), v \in V$$
+% $$min(Q(V), S(V), M(V)), v \in V$$
 
-It should be noted that for practical application this problem can be solved using the constrained optimization method.
+% It should be noted that for practical application this problem can be solved using the constrained optimization method.
 But in practice there are quite strict limitations on the size of the model and the available application time (TODO - what did you mean?).

### 2017-11-11 acat_paper/paper.tex
Added motivation about LHCb

@@
 However it is not always tradeoff between quality vs speed.
-In this paper we show it is possible to speed up model by using additional memory without loosing prediction quality significantly for a novel boosted trees algorithm called Catboost.
+In this paper we show it is possible to speed up model by using additional memory without loosing prediction quality significantly for a novel boosted trees algorithm called CatBoost.
 The idea is to combine two approaches: training fewer trees and merging trees into kind of hash maps (DecisionTensors).
@@
 
-In this paper, we consider a modification of the training process of a predictive models by decision tree-based algorithm CatBoost [1].
-But our approach is applicable to other boosting decision tree-based algorithms.
+For example, Topological Trigger in LHCb is one of the places with very strict applying time restrictions.
+In [2] original MatrixNet [1] model allowed to significantly improve quality.
+But to get acceptable applying time the model was coarsed and converted to Bonsai Boosted Decision Tree format [4].
+
+In this paper, we consider a modification of the training process of a predictive models by successor of MatrixNet, decision tree-based algorithm CatBoost [3].
+But our approach is also applicable to other boosting decision tree-based algorithms.
 
@@
 
-CatBoost is the implementation of a tree-based gradient boosting in which oblivious decision trees are used.
+CatBoost is the implementation of a tree-based gradient boosting in which oblivious decision trees are used [3].
 The values of each feature $i$ are discretized into baskets using the boundaries $B_i={b_{i1},...,b_{ip_i}}$, where $p_i$ is the number of boundaries.
@@
 \item Likhomanenko T, Ilten P, Khairullin E, Rogozhnikov A, Ustyuzhanin A, Williams M 2015 {\it LHCb Topological Trigger Reoptimization} (Journal of Physics: Conference Series)
+\item Dorogush A V, Gulin A, Gusev G, Kazeev N, Ostroumova Prokhorenkova L, Vorobev A {\it Fighting biases with dynamic boosting.} ([arXiv:1706.09516])
+\item Gligorov V, Williams M 2013 {\it Efficient, reliable and fast high-level triggering using a bonsai boosted decision tree} (JINST 8 P02013 [arXiv:1210.6861])
 \endnumrefs

### 2017-11-11 acat_paper/paper.tex
fix
refs

@@
  and
- A. Ustyuzhanin$^{1, 2, 3}$
+ A. Ustyuzhanin$^{2, 3}$
 }
-% \address{Production Editor, \jpcs, \iopp, Dirac House, Temple Back, Bristol BS1~6BE, UK}
-\address{$^{1}$ Moscow Institute of Physics and Technology}
+\address{$^{1}$ Moscow Institute of Physics and Technology, Institutskiy per. 9, Dolgoprudny, Moscow Region, 141700, Russia}
 \address{$^{2}$ National Research University Higher School of Economics, 20 Myasnitskaya st., Moscow 101000, Russia}
@@
 
-\ead{mikari@yandex-team.ru, anaderi@yandex-team.ru}
+\ead{mikari@yandex-team.ru}
 
@@
 Suppose, we have two tensors: $D^1 = (B^{1}_{1}, ..., B^{1}_{n})$ and $D^2 = (B^{2}_{1}, ..., B^{2}_{n})$.
-Their sum $D^u = D^1 + D^2$ could be defined as $D^u = (f^{u}_{1}, ..., f^{u}_{f})$, where $f^{u}_{i} = f^{1}_{i} \cup f^{2}_{i}$.
+The sum $D^u = D^1 + D^2$ could be defined as $D^u = (f^{u}_{1}, ..., f^{u}_{f})$, where $f^{u}_{i} = f^{1}_{i} \cup f^{2}_{i}$.
 Now, let's define similarity as the difference between the sum of their sizes and the size of their union:
@@
 \numrefs{99}
-\item Gulin A, Kuralenok I, Pavlov D 2011 {\it Winning The Transfer Learning Track of Yahoo!'s Learning To Rank Challenge with YetiRank.} (Yahoo! Learning to Rank Challenge) p 63--76
-\item Likhomanenko T, Ilten P, Khairullin E, Rogozhnikov A, Ustyuzhanin A, Williams M 2015 {\it LHCb Topological Trigger Reoptimization} (Journal of Physics: Conference Series)
-\item Dorogush A V, Gulin A, Gusev G, Kazeev N, Ostroumova Prokhorenkova L, Vorobev A {\it Fighting biases with dynamic boosting.} ([arXiv:1706.09516])
-\item Gligorov V, Williams M 2013 {\it Efficient, reliable and fast high-level triggering using a bonsai boosted decision tree} (JINST 8 P02013 [arXiv:1210.6861])
+\item Gulin A, Kuralenok I, Pavlov D 2011 Winning The Transfer Learning Track of Yahoo!'s Learning To Rank Challenge with YetiRank {\it Yahoo! Learning to Rank Challenge} p 63--76
+\item Likhomanenko T, Ilten P, Khairullin E, Rogozhnikov A, Ustyuzhanin A, Williams M 2015 LHCb Topological Trigger Reoptimization (IOP Publishing) {\it Journal of Physics: Conference Series} Vol.
+664, No.
+8, p.
+082025
+\item Dorogush A, Gulin A Gusev G Kazeev N Ostroumova L Vorobev A Fighting biases with dynamic boosting arXiv:1706.09516
+\item Gligorov V, Williams M 2013 Efficient, reliable and fast high-level triggering using a bonsai boosted decision tree {\it Journal of Instrumentation} 8 P02013
 \endnumrefs

### 2018-02-10 acat_paper/paper.tex
Fixed

@@
 \begin{document}
-\title{Speeding up prediction performance of the BDT-based models}
+\title{Speeding up prediction performance of BDT-based models}
 
@@
 The outcome of a machine learning algorithm is a prediction model.
-It is quite usual such models are computationaly expensive and growth of the quality of such models leads to the deterioration in the inference speed.
-However it is not always tradeoff between quality vs speed.
-In this paper we show it is possible to speed up model by using additional memory without loosing prediction quality significantly for a novel boosted trees algorithm called CatBoost.
-The idea is to combine two approaches: training fewer trees and merging trees into kind of hash maps (DecisionTensors).
+Typically, these models are computationally expensive, where improving of the quality the prediction leads to a decrease in the inference speed.
+However it is not always tradeoff between quality and speed.
+In this paper we show it is possible to speed up the model by using additional memory without losing significat prediction quality for a novel boosted trees algorithm called CatBoost.
+The idea is to combine two approaches: training fewer trees and merging trees into a kind of hashmaps called DecisionTensors.
 The proposed method allows for pareto-optimal reduction of the computational complexity of the decision tree model with regard to the quality of the model.
-In the considered example number of lookups was decreased from 5000 to only 6 (speedup factor of 1000) while AUC score of the model was reduced by less than per mil.
+In the considered example the number of lookups was decreased from 5000 to only 6 (speedup factor of 1000) while AUC score of the model was reduced by less than 1$0^{-3}$.
 \end{abstract}
@@
 \section{Introduction}
-Today, Machine Learning (ML) approach is used in various fields ranging from the information retrieval in the Internet and speech recognition to optimization of the steel composition and the search for the New Physics.
-The resources (memory, CPU, network etc.) are very limited to solve the task in production settings.
+Today, Machine Learning (ML) approaches are used in various fields ranging from the information retrieval in the Internet, to speech recognition, to optimizatimg steel compositions, to the search for the New Physics.
+The resources (memory, CPU, network etc.) can be very limited when solving tasks in production settings.
 
-For the most of the ML algorithms, improving the quality of the model usually leads to a higher levels of resource consumption.
-For example, when using tree-based boosting algorithms, the increase of the number of trees from tens to thousands usually leads to significant increase in the quality and at the same time the model size and inference time also grow.
-Higher demands for the computational resources of the more accurate models are usually solved by various technical optimizations model storage mechanism.
+For most ML algorithms, improving the quality of the model usually leads to higher levels of resource consumption.
+For example, when using tree-based boosting algorithms, the increase in the number of trees from tens to thousands usually leads to significant increases in the quality, the model size and inference time.
+Higher demands for the computational resources of the more accurate models are usually satisfied by various technical optimizations of the model storage mechanism.
 For example vectorisation of the tree weights storage can be used.
-Another approach is the modification of the training process that makes it possible to obtain models with certain properties that will provide acceleration to their application without maintaining the required level of quality.
+Another approach is to modify the training process so that it creates models with certain properties that will accelerate their application without maintaining the required level of quality.
 
-For example, Topological Trigger in LHCb is one of the places with very strict applying time restrictions.
-In [2] original MatrixNet [1] model allowed to significantly improve quality.
-But to get acceptable applying time the model was coarsed and converted to Bonsai Boosted Decision Tree format [4].
+For example, the Topological Trigger in the CERN-based LHCb experiment has very strict application time restrictions.
+In [2], the original MatrixNet [1] model allowed for significant improvements in quality.
+However, to get acceptable applying time the model was coarsed and converted to a Bonsai Boosted Decision Tree format [4].
 
-In this paper, we consider a modification of the training process of a predictive models by successor of MatrixNet, decision tree-based algorithm CatBoost [3].
-But our approach is also applicable to other boosting decision tree-based algorithms.
+In this paper, we consider a modification to the training process in the decision tree-based algorithm CatBoost.
+CatBoost is a successor of MatrixNet [3].
+Our approach is also applicable to other boosted decision tree-based algorithms.
 
@@
 The boundaries themselves are determined in advance using various approaches.
-Original feature vector with real values is replaced by a vector of binary values $0$ and $1$ ($false$ and $true$).
-The value of the $i$-th attribute $f_i$ is replaced by a binary vector $G_i={g_{i1},...,g_{ip_i}}$ where $g_{ij} = f_i > b_{ij}$.
+The original feature vector with real values is replaced by a vector of binary values $0$ and $1$ ($false$ and $true$).
+The value of the $i$-th attribute $f_i$ is replaced with a binary vector $G_i={g_{i1},...,g_{ip_i}}$ where $g_{ij} = f_i > b_{ij}$.
 Finally, all $G_i$ are concatenated into one vector.
-Consecutively, the resuling decision will a make bunch of boolean checks instead of comparing some real feature value with the thresholds.
+Consecutively, the resulting decision will a make bunch of boolean checks instead of comparing some real feature value with the thresholds.
 
-Number of baskets used for the discretization could be controlled by specifying the appropriate parameter during the training procedure.
+The number of baskets used for the discretization can be controlled by specifying the appropriate parameter during the training procedure.
 
-Let's consider another important training characteristic, namely the learning rate.
+The learning rate is another important training characteristic in boosted decision-tree methods.
 The smaller it is, the smaller the contribution of each next tree in the ensemble.
@@
 
-With default parameters CatBoost shows already good enough quality, so we variate only number of trees (and learning rate to compensate) because number of trees affects time.
-For other parameters was choosen default values.
+CatBoost shows already acceptable quality with default parameters, so we variate only number of trees (and learning rate to compensate) because number of trees affects time.
+The other parameters were left at their default values.
 
 \section{Decision Tensor}
+
 One of the simplest ways to accelerate computational performance is the preliminary calculation.
 The feature vector obtained after the discretization takes a limited number of values.
-Therefore, we can precompute the predictions for all possible values combination.
+Therefore, we can precompute the predictions for all possible value combinations.
 
@@
 In fact, the model $Z$ allows you to calculate the value at any point of our hypercube.
-if we pre-calculate the value for each piece, then we will accelerate the application of our model, effectively reducing it to the time required to search a piece that contains the point.
-We called such approach a Decision Tensor.
-The asymptotic complexity of the application of the tree-based mode it $\mathcal(h\cdot t)$ where $h$ is the height of the trees, $t$ is the total number of the trees.
-Asymptotic complexity of Tensor is $\mathcal(n\cdot \log(b))$ where $n$ is the number of the features and $b$ is the number of cuts per feature (this complexity is estimated by using binary search).
-It should be noted for small $b$ simple linear search could be a bit faster that binary.
-So, with a small number of features the Tensor-based prediction may be computed much faster than tree-based model.
+If we pre-calculate the value for each piece, then we will accelerate the application of our model, effectively reducing it to the time required to search a piece that contains the point.
+We called this approach a Decision Tensor.
+The asymptotic complexity of the application of the tree-based mode is $\mathcal(h\cdot t)$ where $h$ is the height of the trees, $t$ is the total number of the trees.
+Asymptotic complexity of Tensor is $\mathcal(n\cdot \log(b))$ where $n$ is the number of the features and $b$ is the number of cuts per feature (this complexity is estimated using a binary search).
+It should be noted for a small $b$ a simple linear search could be slightly faster that a binary search.
+With a small number of features the tensor-based prediction may be computed much faster than the tree-based model.
 
-However, Decision Tensor has larger memory footprint.
+However, Decision Tensors have a larger memory footprint.
 It needs $O(b^n)$ bytes of memory and it may be too large for common computer.
@@
 Consider the case for small $n$.
-We can get a small $b$ in two main ways: limit the number of cuts for each feature and train the model with big enough tree count or severely limit the tree count, then the total number of effective cuts will be exactly $h\cdot t$ without taking into account repeated use of the same cuts.
+We can get a small $b$ in two main ways: by limiting the number of cuts for each feature and training the model with a large enough tree count or by severely limiting the tree count.
+In the latter case, the total number of effective cuts will be exactly $h\cdot t$ without taking into account repeated use of the same cuts.
 
-The first method was used in [2]: there a trained MatrixNet (ancestor of CatBoost) model with a small number of cuts that has been precomputed into a Decision Tensor.
-
+The first method was can be found in [2]: which used a trained MatrixNet (ancestor of CatBoost) model with a small number of cuts that has been precomputed into a Decision Tensor.
 The consumption of memory for a single tensor can be extremely huge provided that acceptable quality is achieved, so the natural extension of the idea is the construction of Decision Tensors ensemble but with some deterioration of applying time.
@@
 \section{Algorithm for combining set of Decision Trees to a set of Decision Tensors}
-
-Suppose we have $T$ trees and we want to combine them in $M$ Decision Tensors. The efficiency metric of the partition is the result Tensor set memory footprint.
+Suppose we have $T$ trees and we want to combine them into $M$ Decision Tensors. The efficiency metric of the partition is the resulting memory footprint of the tensor set.
 In general, this is not an easy task and it requires a separate study. We used the following heuristic approach:
@@
 \begin{itemize}
-\item We fix a very rough (with very few ) partition by a small number of Tensors and train an arbitrarily large number of trees.
+\item We use a rough partitioning by a small number of cuts and train an arbitrarily large number of trees.
 \item The partitioning is done quite detailed, but at the same time we significantly reduce the number of trained trees, thereby reducing the number of actually used buckets.
@@
 \item Determine time and size constraints
-\item Determine the desired number of Tensors (M) based on time constraints
+\item Determine the desired number of tensors (M) based on time constraints
 \item Train the model with a limited number of buckets
 \item Merge the trees from the obtained model into the first Decision Tensor from the ensemble
-\item Choose another set of hyperparameters using random search
+\item Choose another set of hyperparameters using a random search
 \medskip
 \begin{itemize}
-\item Train the model, using the previously trained model as the baseline
+\item Train the model, using the previously trained model as a baseline
 \item Transform it into M-1 Decision Tensors
@@
 
-This algorithm allows to find the Pareto-optimal ensemble of Decision Tensors.
+This algorithm identifies the Pareto-optimal ensemble of Decision Tensors.
 
@@
 Number of features in the dataset is 10, the model was trained with 5000 trees and 64 buckets.
-We compared our algorithm for constructing Decision Tensors Ensemble with trivial approach: training a small number of trees but with high value of the training speed.
+We compared our algorithm for constructing Decision Tensors Ensemble with the trivial approach: training a small number of trees but with high value of the training speed.
 For this we divided the data set into three parts: training (50\%), test (25\%) and validation (25\%).
@@
 
-We have limited the number of tensor by 6 and the total size of models by one gigabyte, so every prediction requires 4 bytes (float).
-We have scanned only through the key CatBoost hyperparameters: number of trees and the learing rate.
+We have limited the number of tensors to 6 and the total size of the models to one gigabyte, so every prediction requires 4 bytes (float).
+We have scanned only through the key CatBoost hyperparameters, i.e.
+number of trees and the learning rate.
 
 The figure of merit we've taken also from [2], which is the area under the ROC curve where False Positive Rate less than 0.05.
-For the convenience, we have normalized this metric so that its maximum value (with an ideal model) is equal to 1.
+For the convenience, we have normalized this metric so that its maximum value (with an ideal model) is equal to one.
 
@@
 
-We compared the CatBoost model with 5000 trees, the fast CatBoost models for 1-125 trees and the Decision Tensors Ensembles (DTE) with different memory limits and different tensors numbers.
+We compared the CatBoost model with 5000 trees, the fast CatBoost models for 1-125 trees and the Decision Tensors Ensembles (DTE) with different memory limits and different numbers of tensors.
 
 Figure 1 shows the dependence of the quality on the number of lookups.
-Figure 2 shows the dependence of the quality vs the total inference time at the full validation dataset.
+Figure 2 shows the quality versus the total inference time at the full validation dataset.
 Table 1 shows properties of the models trained in the process.
@@
 This is due to the principles of the CPU caching and the size of our dataset.
-Since the Decision Tensor occupies significant amount of memory and does not fit into the CPU cache almost every lookup results in (slow) reading from the RAM.
+Since the Decision Tensor occupies significant amount of memory and does not fit into the CPU cache almost every lookup results in reading from the RAM, which bottlenecks the working time.
 CatBoost models are small enough to fit into the cache.
@@
 
-Decision Tensor Ensbles allow you to speed up the inference by complex Boosting Decision Trees algorithms (such as CatBoost) in real-time systems.
-Considered algorithm for building Tensor ensembles allows for additional tradeoff between memory and the inference speed without significant loss of the model quality.
-In addition, the use of the Ensemble can significantly improve the model quality in comparison with a single Decision Tensor and can be used in the triggers.
+Decision Tensor Ensembles allow you to speed up the inference by complex Boosted Decision Trees algorithms (such as CatBoost) in real-time systems.
+The algorithm considered for building tensor ensembles allows for additional tradeoff between memory and the inference speed without significant loss of the model quality.
+In addition, the use of the Ensemble can significantly improve the model quality in comparison with a single Decision Tensor and can be used in the LHCb experiment triggers.
 
@@
 082025
-\item Dorogush A, Gulin A Gusev G Kazeev N Ostroumova L Vorobev A Fighting biases with dynamic boosting arXiv:1706.09516
+\item Dorogush A, Gulin A, Gusev G, Kazeev N, Ostroumova L, Vorobev A Fighting biases with dynamic boosting arXiv:1706.09516
 \item Gligorov V, Williams M 2013 Efficient, reliable and fast high-level triggering using a bonsai boosted decision tree {\it Journal of Instrumentation} 8 P02013
